### 2014-03-01 paper/extensions.tex
note on non-deterministic schedulers

\subsection{Safe scheduling policies}
Our specification language is parametrized by a scheduling policy
|alpha|, which maps a task list to a potentially different task list.
%
As previously discussed, the precise definition of |alpha|
dictates whether the language is TSNI, TINI, or neither.
%
For simplicity we considered two popular policies |roundrobinf| and
|seqf| for which we showed TSNI and TINI, respectively.
%
However, our definitions allow for other scheduling policies that only
rely on the current task list.
%
For example, this allows us to implement the scheduler of~\tocite{}
that always schedules less sensitive threads first; this can be
employed to address external timing covert channels where an attacker
can measure the delay between output events.\footnote{
  When considering a label lattice that is not totally ordered, e.g.,
  DCLabels~\tocite{}, this scheduling policy is considerably more
  complex. See~\tocite{}.
}
%
More generally, we belove that deterministic schedulers that always
make ``low-progress,'' i.e., if a task whose label is not above the
current task's label will will eventually run, and do not decide which
low tasks to schedule according to data more sensitive than the tasks'
label, i.e., the scheduler does not schedule one public thread over
another according to sensitive data, to be safe.
%
We leave the proof of this claim to future work.
%

Despite allowing a number of practical schedulers to be expressed, our
scheduler parametrization is limited.
%
For instance, since |alpha| is a map on task lists, schedulers
that rely on the global state |iS|, or previous task configurations
cannot be expressed.
%
Though, we do not belove this to be a fundamental limitation, our
current formalization cannot directly be applied to more-powerful, but
safe, scheduling policies, as considered in~\tocite{}.

>>>>

\subsection{Safe scheduling policies}
Our specification language is parametrized by a scheduling policy
|alpha|, which maps a task list to a potentially different task list.
%
As previously discussed, the precise definition of |alpha|
dictates whether the language is TSNI, TINI, or neither.
%
For simplicity we considered two popular policies |roundrobinf| and
|seqf| for which we showed TSNI and TINI, respectively.
%
However, our definitions allow for other scheduling policies that only
rely on the current task list.
%
For example, this allows us to implement the scheduler of~\tocite{}
that always schedules less sensitive threads first; this can be
employed to address external timing covert channels where an attacker
can measure the delay between output events.\footnote{
  When considering a label lattice that is not totally ordered, e.g.,
  DCLabels~\tocite{}, this scheduling policy is considerably more
  complex. See~\tocite{}.
}
%
More generally, we believe that deterministic schedulers that always
make ``low-progress,'' i.e., if a task whose label is not above the
current task's label will will eventually run, and do not decide which
low tasks to schedule according to data more sensitive than the tasks'
label, i.e., the scheduler does not schedule one public thread over
another according to sensitive data, to be safe.
%
We leave the proof of this claim to future work.
%

Despite allowing a number of practical schedulers to be expressed, our
scheduler parametrization is limited.
%
For instance, since |alpha| is a map on task lists, schedulers that
rely on the global state |iS| or previous task configurations cannot
be expressed.
%
Though, we do not believe this to be a fundamental limitation, our
current formalization cannot directly be applied to more-powerful, but
safe, scheduling policies, as considered in~\tocite{}.
%
More fundamentally, our definitions rely on deterministic relations,
and thus extending the system to consider non-deterministic schedulers
is non-trivial.
%
For instance, this may potentially require the security condition to
consider \emph{low-determinism}, which states that a program is secure
only if |l|-equivalent results are deterministic~\tocite{}. %both andrei and steve's papers
